---
title: "Improving Runtime"
author: "Robert M Flight"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

ICI-Kt can take a long time to run.
Subsets can significantly speed it up, but taking a subset that is too small or incorrect will result in non-representative sample-sample correlations.

There are two possible ways to rectify this:

  1. Select features that have a higher variance.
  1. Use PCA to select features with high loadings on each PC.
  
For (2), Hunter has suggested we use the proportion of variance for each PC to determine the total proportion of high-loading features in each PC to use.

Let's work out the logic for this, which can then be run in separate functions.

## Variance

```{r variance_selection}
var_select = function(matrix_data, n_item){
  var_rows = apply(matrix_data, 1, var)
  var_data = data.frame(row_id = rownames(matrix_data),
                        var = var_rows)
  var_data = dplyr::arrange(var_data, dplyr::desc(var))
  use_rows = var_data$row_id[1:n_item]
  matrix_data[use_rows, ]
}
```

## PCA

```{r pca_selection}
pca_select = function(pca_data, matrix_data, n_item){
  pc_cont = visqc_score_contributions(as.matrix(pca_data$x))
  use_pcs = dplyr::filter(pc_cont, cumulative <= 0.95)
  use_pcs$n_row = round(n_item * use_pcs$percent)
  use_rows = vector("character", n_item)
  
  if (is.null(rownames(matrix_data))) {
    tmp_names = seq(1, nrow(matrix_data))
  } else {
    tmp_names = rownames(matrix_data)
  }
  
  count_load = 1
  use_pc = 1
  while (count_load < n_item) {
    use_loadings = pca_data$rotation[, use_pc]
    names(use_loadings) = tmp_names
    use_loadings = sort(abs(use_loadings), decreasing = TRUE)
    use_loadings = use_loadings[!(names(use_loadings) %in% use_rows)]
    use_loc = seq(count_load, count_load + use_pcs$n_row[use_pc] - 1)
    use_loc = use_loc[use_loc <= n_item]
    load_loc = seq(1, length(use_loc))
    use_rows[use_loc] = names(use_loadings[load_loc])
    count_load = use_loc[length(use_loc)] + 1
  }
  
}
```
