---
title: "Similarity Tables"
author: "Robert M Flight"
output: 
  rmarkdown::word_document:
    keep_md: true
date: '`r Sys.time()`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
targets::tar_source("./packages.R")
tar_source("R")

figure_count = dn_counter$new("Figure ", "_")
table_count = dn_counter$new("Table ", "_")

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 8, 
                      fig.height = 6, 
                      fig.process = figure_count$modify_path,
                      dpi = 600,
                      dev = "ragg_png")
```

```{r}
#| label: load-dependencies
tar_load(c(feature_compare_summary
))

```

```{r}
#| label: split-table
split_summary = split(feature_compare_summary, feature_compare_summary$data_id)
all_tables = purrr::map(split_summary, \(x){
  x |>
    dplyr::select(-data_id) |>
    cleanup_significant_compare_table()
})
```

Table X. Yeast significant edge comparisons.
`n1` is number of edges in set 1, `n2` is number of edges in set 2, `min` is the smaller number of edges, `jaccard` is the Jaccard index, `overlap` is the Overlap index, and `combined` is essentially an average of the two.
`d1` is dataset 1, `d2` is dataset 2.

```{r}
all_tables$barton_yeast
```

Table X. EGFR Genotype significant edge comparisons.
`n1` is number of edges in set 1, `n2` is number of edges in set 2, `min` is the smaller number of edges, `jaccard` is the Jaccard index, `overlap` is the Overlap index, and `combined` is essentially an average of the two.
`d1` is dataset 1, `d2` is dataset 2.

```{r}
all_tables$brainsonrnaseq_egfrgenotype
```

Table X. NSCLC significant edge comparisons.
`n1` is number of edges in set 1, `n2` is number of edges in set 2, `min` is the smaller number of edges, `jaccard` is the Jaccard index, `overlap` is the Overlap index, and `combined` is essentially an average of the two.
`d1` is dataset 1, `d2` is dataset 2.

```{r}
all_tables$nsclc
```

Table X. Rat Stamina significant edge comparisons.
`n1` is number of edges in set 1, `n2` is number of edges in set 2, `min` is the smaller number of edges, `jaccard` is the Jaccard index, `overlap` is the Overlap index, and `combined` is essentially an average of the two.
`d1` is dataset 1, `d2` is dataset 2.

```{r}
all_tables$mwtab_ratstamina
```
