---
title: 'Supplemental Tables'
output: 
  rmarkdown::word_document:
    reference_docx: 'table_template.docx'
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, 
                      message = FALSE, fig.width = 8, 
                      fig.height = 6, fig.keep = "all",
                      fig.process = dn_modify_path,
                      tab.cap.style = "paragraph",
                      tab.cap.pre = "",
                      tab.cap.sep = "",
                      dpi = 600,
                      dev.args = list(png = list(type = "cairo")))
tar_load(supp_materials)
supp_figure_count = readRDS(here::here("doc/supp_figure_count.rds"))
supp_table_count = readRDS(here::here("doc/supp_table_count.rds"))
load(here::here("doc/supp_stuff.rda"))
```

`r supp_table_count$label_text("yeast_outliers")`.
Yeast dataset median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

```{r yeast_table2}
#yeast_ft_out = set_caption(yeast_table,
  # caption = paste0(paste_label("Table S", supp_table_count, "yeast_outliers"), ". Yeast dataset median correlation values and outlier determination for each outlier from each of the correlation methods."), style = "paragraph")
yeast_ft_out = yeast_table %>%
  set_nice_widths(sample_width = 0.52, outlier_width = 0.5, cor_width = 0.7) %>%
  put_outlier_right() %>%
  fontsize(size = 7, part = "all")
yeast_ft_out
```


`r supp_table_count$label_text("egfrgenotype_outliers")`.
EGFR RNA-Seq dataset using genotype as the grouping variable median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures and data are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

```{r egfrgenotype_table2}
egfrgenotype_ft_out = egfrgenotype_table %>%
  set_nice_widths(sample_width = 1.1, outlier_width = 0.5, cor_width = 0.7) %>%
  put_outlier_right() %>%
  fontsize(size = 7, part = "all")
egfrgenotype_ft_out
```

`r supp_table_count$label_text("egfrgenotypetumorculture_outliers")`.
EGFR RNA-Seq dataset using growth method as the grouping variable median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures and data are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

`r supp_table_count$label_text("adenocarcinoma_outliers")`.
TGCA adenocarcinoma dataset median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures and data are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

```{r adenocarcinoma_table2}
adenocarcinoma_ft_out = adenocarcinoma_table %>%
  set_nice_widths(sample_width = 0.52, outlier_width = 0.5, cor_width = 0.7) %>%
  put_outlier_right() %>%
  fontsize(size = 7, part = "all")
adenocarcinoma_ft_out
```

`r supp_table_count$label_text("nsclc_outliers")`.
NSCLC lipidomics dataset median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures and data are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

```{r nsclc_table2}
nsclc_ft_out = nsclc_table %>%
  set_nice_widths(sample_width = 0.52, outlier_width = 0.5, cor_width = 0.7) %>%
  put_outlier_right() %>%
  fontsize(size = 7, part = "all")
nsclc_ft_out
```

`r supp_table_count$label_text("ratstamina_outliers")`.
Rat stamina metabolomics dataset median correlation values and outlier determination for each outlier from each of the correlation methods.
Abbreviations for different measures and data are: IK: ICI-Kt; IKC: ICI-Kt * Completeness; PB: Pearson Base, PN: Pearson No Zero; PL1: Pearson Log(x + 1); PL: Pearson Log(x); Kt: Kendall-tau.

```{r ratstamina_table2}
ratstamina_ft_out = ratstamina_table %>%
  set_nice_widths(sample_width = 0.52, outlier_width = 0.5, cor_width = 0.7) %>%
  put_outlier_right() %>%
  fontsize(size = 7, part = "all")
ratstamina_ft_out
```
